Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-29 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

While the usefulness of deep audio embeddings can be evaluated in several scenarios, we have chosen a core MIR task for its popularity and complexity: Music Structure Analysis (MSA). This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

SALAMI (Structural Analysis of Large Amounts of Music Information) \cite{Smith2011DESIGNANNOTATIONS} aims to conduct massive structural analyses of different types of music. The project covers diverse music, from Western pop to Indian classical and from live to studio recordings.

\subsubsection{Formal analysis definition and approaches}
Formal analysis, in this context, doesn't refer to classifying the piece into a formal type (e.g., sonata, song, or canon) or reducing it into its Ursatz as in Schenkerian analysis. Instead, it means organizing and dividing the piece into specific sections and understanding how they relate.

The proposed method integrates aspects of three approaches: perceptual, functional, and transcription, enabling a comprehensive analysis of a wide range of music pieces using a limited vocabulary.

It considers the hierarchical nature of music structure and employs markers to partially address the challenge of determining the appropriate timescale for analysis.

The perceptual approach involves segmenting and labeling sections based on harmonic or rhythmic boundaries, although it may not accurately reflect the structural roles of similar-sounding sections.

The functional approach categorizes the song into components like verses and choruses, potentially leading to varied labels for identical music or identical labels for different musical ideas based on their function.

The transcription approach records musical parameters directly, such as chord patterns and instrumentation, providing a less subjective method, but it might not capture the complete structure of the piece.

@@

The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

Below is a detailed explanation of computing the Embeddiogram from a given audio signal. This process comprises five key steps, including loading the audio data, slicing the audio data into windowed segments, processing each window using a pre-trained model to produce an embedding, collecting these embeddings, and finally normalizing the Embeddiogram.

@@

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{Embeddiogram}, denoted as $\text{embeddiogram}$:
\begin{equation}
\text{embeddiogram} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the Embeddiogram}: The Embeddiogram is normalized to have a minimum value of 0 and a maximum value of 1. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/embeddiogram SALAMI 3.png}
    \caption[Embeddiogram]{Embeddiogram computed from...}
    \label{fig:embeddiogram}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/noveltypeaks_SALAMI_track_2.png}
    \caption[Novelty curve]{Caption}
    \label{fig:embeddiogram}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/recurrence plot SALAMI 3.png} % first figure itself
        \caption{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/Rec_matrix_smooth_SALAMI_track_2.png} % second figure itself
        \caption{second figure}
    \end{minipage}
\end{figure}

>>>>

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks, we have chosen Music Structure Analysis (MSA) for its popularity, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

SALAMI (Structural Analysis of Large Amounts of Music Information) aims to conduct massive structural analyses of different types of music \cite{Smith2011DESIGNANNOTATIONS}. The project covers diverse music, from Western pop to Indian classical and from live to studio recordings.

The authors propose a unique music analysis approach that segments pieces into sections rather than traditional forms. Their method combines perceptual, functional, and transcription analyses, considers music's hierarchical nature, and uses markers to determine analysis timescales. While providing a comprehensive investigation, each analysis approach has limitations in accurately reflecting musical structure, labeling identical or different music sections, and capturing the complete structural framework. This innovative approach aims for a more nuanced, holistic music analysis.

@@

The \textit{embeddiogram} is derived by applying our pre-trained neural network model to sliding windowed segments across the audio signal, generating a sequence of embedding vectors. These relatively high-dimensional embedding vectors collectively form a two-dimensional description of the audio signal's musical content.

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal. This process comprises five key steps, including loading the audio data, slicing the audio data into windowed segments, processing each window using a pre-trained model to produce an embedding, collecting these embeddings, and finally normalizing the Embeddiogram.

@@

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{embeddiogram}, denoted as $\text{e}$:
\begin{equation}
\text{e} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the \textit{embeddiogram}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{.8\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/salami_391_embeddiogram.png}
    \caption[]{Caption for image1}
    \label{fig:image1}
  \end{minipage}

  \begin{minipage}[b]{.8\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/samali_391_novelty_curve_peaks.png}
    \caption[]{Caption for image2}
    \label{fig:image2}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_R matrix after Gaussian smoothing.png} % first figure itself
        \caption[]{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_Lag matrix after Gaussian smoothing.png} % second figure itself
        \caption[]{second figure}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_Q_matrix.png} % third figure itself
        \caption[]{third figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_binary transitive matrix.png} % fourth figure itself
        \caption[]{fourth figure}
    \end{minipage}
\end{figure}

### 2023-05-29 future work/future work.tex
Updates from Overleaf

hybrid autoencoder architecture + UNet stuff
transforms
stems
different takes

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; are a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies. In addition to their wide usage, it's been recently reported that their frequency-domain summarization, grounded in psycho-acoustics, is efficient and challenging to replicate through solely data-driven methods. \cite{Kim2020OneStrategies}

>>>>

hybrid autoencoder architecture + UNet stuff
transforms
stems
different takes
k-fold cross validation

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. In addition to their wide usage, it's been recently reported that their frequency-domain summarization, grounded in psycho-acoustics, is efficient and challenging to replicate through solely data-driven methods. \cite{Kim2020OneStrategies}

\newpage

### 2023-05-29 methods/subsections/implementation.tex
Updates from Overleaf

\subsection{Contrastive Learning of Musical Representations (CLMR)}

@@

The CLMR methodology involves the following steps:

\begin{enumerate}
\item \textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as antagonistic pairs.
\item \textbf{Contrastive Learning:} Train by minimizing a contrastive loss function, encouraging the model to produce similar representations for positive pairs and dissimilar representations for antagonistic pairs.
\end{enumerate}

\subsection{Siamese and Triplet Siamese Networks}

@@

The triplet loss function used in this network guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space. Due to the shared weights among the subnetworks, Triplet Siamese Networks also offer efficient memory usage and computational complexity.

@@

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

@@

%%%%%BATCH NORM%%%%%%%%%%

In implementing batch normalization, it is necessary to standardize the audio lengths across all elements in the minibatch. We opted to zero-pad all clips to the length of the longest clip, valuing data integrity and completeness over potential performance trade-offs. Thus, the length of the longest array in the batch, which sets the standard for all others, is as follows:

\begin{equation}
L_{\text{max}} = \max_{\text{item} \in \text{batch}}(\max(\text{length}(A_{\text{item}}), \text{length}(P_{\text{item}}), \text{length}(N_{\text{item}})))
\end{equation}

@@

@@

The positive image in our study should preserve its intelligible content when subjected to temporal transformations, regardless of alterations in sonic qualities. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or strikingly similar to the original track.

While the MIR community has access to helpful audio transformation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0}, the specific requirements of our experiments necessitated the development of our transformation chain: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

@@

@@

In this equation, the triplet loss function with a variable margin:

$\mathcal{L}$ is the triplet loss function.
$\mathbf{a}_i$, $\mathbf{p}_i$, and $\mathbf{n}_i$ are the anchor, positive, and negative embedding vectors, respectively, for the $i$-th sample.
$\left| \cdot \right|_2^2$ denotes the squared Euclidean distance between two points.
$m$ is the margin, a hyperparameter that helps ensure that the anchor-positive distance is smaller than the anchor-negative distance by at least a margin.
The summation $\sum_{i=1}^{N}$ is over all $N$ triplets in the dataset.
The goal of the triplet loss is to minimize the distance between the anchor and the positive sample while maximizing the distance between the anchor and the negative sample. This encourages the model to learn embeddings that are similar for the same class and dissimilar for different classes.

The final form of the loss function is obtained by averaging the loss over a mini-batch of triplets.

\begin{equation}
\mathcal{L} = \frac{1}{N} \sum_{i=1}^{N} L(anchor_i, positive_i, negative_i)
\end{equation}

Here, $N$ is the number of triplets in the mini-batch, and $L(anchor_i, positive_i, negative_i)$ represents the loss function computed for the i-th triplet in the batch.

\subsubsection{Triplet mining}

The anchor-positive and anchor-negative distances are computed as the Euclidean distances between the corresponding samples.

\begin{equation}
D_{\text{AP}} = \sqrt{\sum_{i} (A_i - P_i)^2}
\end{equation}

The "hardest" negative index for each anchor-positive pair is determined by maximizing the difference between the anchor-negative and anchor-positive distances \cite{XuanImprovedMining}.

\begin{equation}
D_{\text{AN}} = \sqrt{\sum_{i} (A_i - N_i)^2}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[clip,width=\columnwidth]{figures/images/Triplet Network.jpeg}
    \caption[]{Caption}
    \label{fig:my_label}
\end{figure}

>>>>

\section{Implementation details}
\subsection{Contrastive Learning of Musical Representations}

@@

The CLMR methodology involves creating different augmentations of the same musical piece as positive pairs and using others as antagonistic pairs. Then, training by minimizing a contrastive loss function encourages the model to produce similar representations for positive pairs and dissimilar representations for antagonistic pairs.

\subsection{Triplet Siamese Networks}

@@

@@

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes 1D input data with a single channel and follows the next architecture:

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c}{The following layers are repeated depending on the strides and hidden parameters:} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected (representation layer) & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnn}
\end{table}

@@

\subsubsection{Optimizer}

@@

We believe our experiments can benefit from time-domain processing because it allows more natural handling of temporal patterns and sequences in the data while avoiding windowing artifacts.

Mel-spectrogram-based approaches have proved useful in numerous audio-related machine-learning tasks using CNN, as their frequency-domain summarization is efficient and not easily learnable through data-driven approaches \cite{Kim2020OneStrategies}. That being said,  they exhibit limitations concerning their representation of perceptual similarity: the spectrogram of an audio signal, denoted as $\phi(x)$ for audio signal $x$ and $\phi(y)$ for audio signal $y$, captures the distribution of frequencies as they vary over time. A small difference between $\phi(x)$ and $\phi(y)$ generally signifies high perceptual similarity from a human listener's perspective. In other words, if two audio signals have similar spectrograms, they are likely to sound alike when heard.

However, the inverse is not consistently true. There exist instances where two audio signals can have significantly different spectrograms, i.e., a high spectrogram distance between $\phi(x)$ and $\phi(y)$, but are still perceived as musically similar by human listeners. This discrepancy is due to the complexity of human auditory perception that goes beyond just the distribution of frequencies over time \cite{Mesostructures2023}.

Our perception of sound involves several other factors, including temporal patterns, phase relationships between frequencies, and even higher-level musical structures. Therefore, even if two sounds have very different spectrograms, they might still be perceived as similar if they share these other characteristics. This underlines the inherent limitation of using only spectrogram distance to measure high-level music similarity, as it may not always produce results consistent with human perceptual judgments.

@@

The positive image in our study must preserve its intelligible content when subjected to temporal (or any) transformations, regardless of alterations in sonic qualities. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or strikingly similar to the original track.

While we have experimented with helpful audio augmentation tools such as \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0} and \cite{Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our transformation chain using \textit{torchaudio} \cite{Yang2021TorchAudio:Processing}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

@@

The triplet loss function used in Siamese networks guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

The triplet loss function $\mathcal{L}(\mathbf{a}, \mathbf{p}, \mathbf{n})$ aims to ensure that an anchor vector $\mathbf{a}_i$ is closer in the embedding space to a positive vector $\mathbf{p}_i$ (representing an example of the same class) than to a negative vector $\mathbf{n}_i$ (representing an example of a different class) by at least a margin $m$. It is calculated by summing the losses overall $N$ triplets in the dataset, where the equation gives the loss for each triplet:

@@

The final loss used for model training is then the average loss over a mini-batch of $N$ triplets:

\begin{equation}
\mathcal{L} = \frac{1}{N} \sum_{i=1}^{N} \mathcal{L}(\mathbf{a}_i, \mathbf{p}_i, \mathbf{n}_i)
\end{equation}

The goal of minimizing this loss function is to learn discriminative embeddings, where similar examples are grouped closely together and dissimilar examples are placed farther apart in the embedding space.

\subsection{Online triplet mining and batch normalization}

Online triplet mining is beneficial for managing large datasets by dynamically selecting the most informative triplets during training, focusing on each mini-batch. This strategy makes the process memory-efficient by negating the need to store all possible triplet combinations. Still, it also enhances model performance through quicker convergence by focusing on challenging examples based on the current model state.

This hard triplet mining selects triplets $(a, p, n)$ to maximize the Euclidean distance between the anchor and positive samples and the anchor and negative samples. These distances, $D_{\text{AP}}$ and $D_{\text{AN}}$, are computed respectively as:

\begin{align}
D_{\text{AP}} &= \sqrt{\sum_{i} (A_i - P_i)^2} & D_{\text{AN}} &= \sqrt{\sum_{i} (A_i - N_i)^2}
\end{align}

\subsubsection{Batch normalization}

In implementing the batch normalization step, it is necessary to standardize the audio lengths across all elements in the minibatch. We opted to zero-pad all clips to the length of the longest clip, valuing data integrity and completeness over potential performance trade-offs. Thus, the length of the longest array in the batch, which sets the standard for all others, is as follows:

\begin{equation}
L_{\text{max}} = \max_{i \in I} \left( \max \left( |A_i|, |P_i|, |N_i| \right) \right)
\end{equation}

\begin{itemize}
    \item $I$ represents the set of all items in the batch.
    \item $|A_i|$, $|P_i|$, and $|N_i|$ denote the lengths of the anchor, positive, and negative embedding vectors for the $i$-th item, respectively.
    \item The $\max$ function is applied to find the longest of these three lengths for each item, and then the maximum of these maximum lengths is taken over all items in the batch. This gives the maximum length, $L_{\text{max}}$, of any vector in the batch.
    \end{itemize}

\newpage

### 2023-05-29 results/results.tex
Updates from Overleaf

>>>>

Extend table by \cite{Hernandez-Olivan2021MusicFeatures}

\begin{table}[h]
\centering
\footnotesize
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|}
\hline
Year & Authors & Ref. & Algorithm & Input & Method & Supervised? & F-measure SALAMI \\
\hline
2009 & Paulus \& Klapuri & [24] & PK & MFCCs, chromas & Fitness function & No & - \\
2010 & Mauch et al. & [25] & MND1 & MFCCs, Discrete Cepstrum & HMM & No & - \\
2011 & Sargent et al. & [26] & SB-VRS1 & Chords estimation & Viterbi & No & - \\
2012 & Kaiser et al. & [27] & KSP2 & SSM & Novelty measure & No & 0.286 \\
2013 & McFee \& Ellis & [20] & MP2 & MLS & Fisher’s Linear Discriminant & No & 0.317 \\
2014 & Nieto \& Bello & [28] & NB1 & MFCCs + chromas & Checkerboard-like kernel & No & 0.299 \\
2015 & Cannam et al. & [29] & CC1 & Timbre-type histograms & HMM & No & 0.213 \\
2016 & Nieto & [30] & ON2 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis & No & 0.299 \\
2017 & Cannam et al. & [29] & CC1 & Timbre-type histograms & HMM & No & 0.212 \\
2014 & Schlüter et al. & [31] & SUG1 & MLS & CNN & Yes & 0.529 \\
2015 & Grill \& Schlüter & [32] & GS1 & MLS + SSLMs & CNN & Yes & 0.541 \\
2007 & Turnbull et al. & [19] & - & MFCCs, chromas, spectrogram & Boosted Decision Stump & No & 0.378 \\
2011 & Sargent et al. & [34] & - & MFCCs, chromas & Viterbi & No & 0.356 \\
2014 & Ullrich et. al & [22] & - & MLS & CNN & No & 0.465 \\
2015 & Grill \& Schlüter & [4] & - & MLS + SSLMs & CNN & No & 0.523 \\
2015 & Grill \& Schlüter & [5] & - & MLS + PCPs + SSLMs & CNN & No & 0.508 \\
2017 & Hadria \& Peeters & [35] & - & MLS + SSLMs & CNN & No & 0.291 \\
\hline
\end{tabularx}
\caption{Your Caption}
\label{tab:my_label}
\end{table}
